InputFinder: Reverse Engineering Closed Binaries via Side Channels

### 2015-05-04 paper/main.tex
updated paper

% definitions
\def \tool {InputFinder}

@@

\section{Methodology} \label{methodology}
% add some high level intro for both finding input and psm stuff
\tool{} is composed of two main modules.
The first module is responsible for discovering input and properties of the expected input, including input size and the type of the input.
The input can then be used by the second component to generate a protocol state machine, information which may then be used by a stateful fuzzer.
\subsection{Finding Valid Input}
The majority of software programs accept input, perform transformations on the input and output results.
However, most programs do not accept completely random input.
Input usually passes through an input validation filter.
Such filters are snippets of code responsible for distinguishing between valid or good input (i.e. input the program was constructed to understand and accept) and bad input (i.e. input which is not useful or does not follow the desired format).
Although validation mechanisms can be very sophisticated, they are often a combination of string comparisons and conditional statements.
Our method exploits changes in the number of user-land instructions retired during multiple executions with varying inputs to make inferences about the program's validation mechanism.
These changes reflect different execution paths of the program as a response to both valid and invalid input.
Assuming most input is invalid, our method observes differences in the number of instructions executed to detect input which passes the validation mechanism.

@@

\subsubsection{Finding Desired Input Size}
The same side channels can be used to detect the expected size of the program's input or even the program's input buffer.
Most programs are designed to read a predefined number of bytes from the source of input and store them in a buffer.
Depending on the implementation, when a program is given more input than the size of the buffer, additional instructions may be executed.
Regardless of the number of instructions executed, the behavior of the program changes.
A program may exit when the number of provided bytes is larger than expected.
Similarly, once the input buffer size limit has been reach, a program may check to see if the next byte is a newline or EOF.
Some methods, such as C's `fgets()' function, read character by character [EXPLAIN]
Our method observes such differences in the number of user-instructions retired as the program is given increasingly larger input.

To start, \tool{} finds the set of all invalid characters for index 0 of all input strings and picks one at random.
The character selected is used to build invalid input strings of various sizes.
It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions executed due to valid input processing.
Initially, the program is executed with short length input strings (i.e. two and three characters).
The number of user-land instructions executed during these two executions are recorded.
These two values may be equal or differ, depending on the program's logic.
If the two values are equal [EQUAL]
The two values can also differ because... [DIFFERENT]

\subsubsection{Categorizing Input}
Our method is also capable of classifying the input discovered.
Generated input strings are classified in three categories: alphabetic, numeric, alphanumeric.
Similarly, the classification algorithm detects special characters.
Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program accepts.
Such information is valuable especially in cases where the inputs discovered are all numeric or contain (one or more) special characters.
Having learned such information about the program, the testing platform (e.g. fuzzer) can be adjusted to proceed accordingly.

@@

Once input has been discovered, \tool{} constructs a protocol state machine.
The architecture of this component is depicted in Figure \ref{fig:XXXXpsm_arch}.
The process is describe below and shown as pseudocode in Listing \ref{lst:psmcode}.
First, all possible permutations of all sizes from 1 to the number of input strings discovered are computed.
Every permutation is passed through the dynamic analysis engine.
The dynamic analysis execution engine first splits the permutation into individual input strings.
Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
For example, if the input string is `SET' and the program accepts input of the format `SET data', the analysis tries to determine that `data' is a valid argument.
The arguments found are saved for later use.
Once every input pertaining to a permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
The binary is instrumented with a Pin tool, developed by us, which logs the instructions executed as a function of time.
Once all permutations of a given size have been executed, the dynamic similarity test begins.
This test splits each execution trace into parts, each part being associated with an input string.
%Next, the execution trace part associated with the last input string in the permutation is transformed into the depth-first search (DFS) string representation of its control flow graph (CFG).
We refer to the execution trace of the input string `SET' as $\alpha(SET)$.
For all permutations of the same size with the same last input string, the $ [last_cmd] $ are clustered.
Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the last input of the permutation experiences a different execution trace (or control flow path) compared to that of the same command in an incorrect permutation (assuming it is the same index in the permutation).
This information is then used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
If such a permutation is discovered, it it added to the protocol state machine.

@@

\section{Evaluation and Results} \label{results}
\subsection{Finding Input, Input Size and Classifying Input}
\tool{} was tested on [NUMBER] closed-source binaries to measure its effectiveness.
The results are summarized in Figure [FIGURE].
Out of the total [NUMBER] binaries, [NUMBER2] had predefined accepted input strings.
\tool{} found 100\% of the input strings for [NUMBER3]...
[CONTINUE].
%To further show the advantage of this method, we built a smart fuzzer which uses our method to determine input and protocol for a given binary.
%We use the smart fuzzer and compare it to black box fuzzers symbolic execution engines with respect to code coverage over units of time.
%Using such a comparison allows the reader to determine which method is more efficient while also performing well.

@@

Additionally, our method was tested on a binary used for technical interviews.
The binary was designed as a reverse engineering challenge for interviewers.
To pass the challenge, the interviewer has to crack the secret passcode to the program.
The blog post describes the manual process used to crack the password \cite{interviewbinary}.
The manual process is tedious and it includes methods such as disassembling the binary, stepping through the instructions using a debugger, and reverse engineering the logic of the program.
As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
\tool{} cracked the binary in a matter of minutes autonomously, proving itself as a useful tool not only finding valid input but also cracking passwords.

\subsection{Protocol State Machine Generation}
% TODO
The evaluation of the protocol state machine generator was also done using the DARPA Cyber Grand Challenge binaries.
However, out of the total of [NUMBER] binaries, only [NUMBER] had an implicit protocol.
Our method was able to generate a correct protocol state machine for [NUMBER]\% of these binaries.

@@

\section{Future Work and Limitations} \label{futurework}
While the method presented in this paper has many advantages, it also has a few limitations.
For certain programs that allow any input, such as an echo server, our method is not any more efficient than a black box fuzzer.
Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field.
Also, \tool{} cannot handle programs which validate input fields out of order or use any safe hashing algorithm for their input validation technique. 
In the future, we would like to apply the input finding technique recursively.
It is easy to imagine programs for which the second input varies based on the value of the first input.
Also, as we've experienced during our evaluation, there exist programs which first request for a username, input which will not have any predefined valid values.
It would be helpful to determine such cases, provide some input and continue the execution of the program, applying the input finding technique once again.
However, there are challenges with this, such as determining when to stop.

With respect to the protocol state machine, currently our method uses the input finding technique to determine valid arguments.
However, the protocol state machine does not currently handle situations in which there are no predefined arguments (e.g. given input ``CREATE'' the arguments could be predefined such as ``dog'' or random such as ``x'').
Ultimately, we would like to generate some random arguments that remain constant for every input string in a given permutation.
Moreover, our current approach is not capable of identifying equivalent states. As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of ``HELLO" vs. omission of ``HELLO") lead to different states when in reality the two states are identical.
Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.
There may be cases where this assumption does not hold.
In the future, we would like build onto our current approach for generating protocol state machines. 
First, we want to be able to determine if two states are identical for a more accurate protocol state machine.
Furthermore, we want to determine a better approach for identifying the valid sequence of input strings.
One option would be to monitor other aspects of the program, particularly the \textit{.data}, \text{.bss}, and the heap during the executions.
Changes in these may reflect changes of the state of the program.

@@

We show that our proposed methods are efficient and successful by testing them on [NUMBER] DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle even binaries of other formats.

>>>>

% definitions
\def \tool {InputFinder}  
\def \numbinaries {24}

@@

\section{Methodology} \label{methodology}

\tool{} is composed of two main modules.
The first module is responsible for discovering input and properties of the expected input, including input size and the type of the input.
The input can then be used by the second component to generate a protocol state machine, information which may then be used by a stateful fuzzer.
\subsection{Finding Valid Input}
The majority of software programs accept input, perform transformations on the input and output results.
However, most programs do not accept completely random input.
Input usually passes through an input validation filter.
Such filters are snippets of code responsible for distinguishing between valid or good input (i.e. input the program was constructed to understand and accept) and bad input (i.e. input which is not useful or does not follow the desired format).
Although validation mechanisms can be very sophisticated, they are often a combination of string comparisons and conditional statements.
Our method exploits changes in the number of user-land instructions retired during multiple executions with varying inputs to make inferences about the program's validation mechanism.
These changes reflect different execution paths of the program as a response to both valid and invalid input.
Assuming most input is invalid, our method observes differences in the number of instructions executed to detect input which passes the validation mechanism.

@@

\subsubsection{Finding Desired Input Size}
The same side channels can be used to detect the expected size of the program's input or even the size of the program's input buffer.
Specifically, there is a change in the behavior of the program with respect to the number of instructions executed when the size of the input is larger than expected.
Most programs are designed to read a predefined number of bytes from the source of input and store the read bytes in a buffer.
A popular way of doing that is by using the \textit{fgets} or \textit{scanf} functions of the GNU C stdio library.
As the input size grows, the number of instructions executed also increases linearly.
This happens because additional instructions are executed to read more characters and store them into the buffer.
However, when the input string is larger than the expected size the number of instructions executed no longer increases as a function of the input size.
This behavior can be observed and exploited to find the expected input size, which may often also be the size of the input buffer.
Our method observes such differences in the number of user-land instructions retired as the program is given increasingly larger input.

To start, \tool{} finds the set of all invalid characters for index 0 of all input strings and picks one at random.
The character selected is used to build invalid input strings of various sizes.
It is crucial to use invalid input to avoid noisy fluctuations in the number of instructions executed due to valid input processing.
Initially, the program is executed with short length input strings (i.e. two and three characters).
The number of user-land instructions executed during these two executions are recorded.
These two values may be equal or differ, depending on the program's logic.
If the two values are equal, the program is executed with increasingly large input until a change in the number of executions is recorded.
%TODO
The two values may be equal if the function responsible for reading input ..
If the values differ (as discussed earlier), the difference in the number of instructions executed is recorded and the program is executed, increasing the size of the input, until a change is detected.
Instead of increasing the input size linearly, the input size is increased exponentially until a change is detected.
Once the change is detected, the input size is decreased to find the exact size at which the change happens.

\subsubsection{Categorizing Input}
Our method is also capable of classifying the input discovered.
Generated input strings are classified in three categories: alphabetic, numeric, alphanumeric.
Similarly, the classification algorithm detects special characters.
Once enough input strings have been identified, \tool{} uses the categorization data to generalize about the type of input the program accepts.
This is done by iterating through the characters of the input strings discovered and determining which category they fall into.
Such information is valuable especially in cases where the inputs discovered are all numeric or contain (one or more) special characters.
Having learned such information about the program, the testing platform (e.g. fuzzer) can be adjusted to proceed accordingly.

@@

Once input has been discovered, \tool{} constructs a protocol state machine.
The architecture of this component is depicted in Figure \ref{fig:psm_arch}.
The process is describe below and shown as pseudocode in Listing \ref{lst:psmcode}.
First, all possible permutations of all sizes from 1 to the number of input strings discovered are computed.
Every permutation is passed through the dynamic analysis engine.
The dynamic analysis execution engine first splits the permutation into individual input strings.
Maintaining the original order of the inputs, the engine tests each string for the presence of expected arguments.
For example, if the input string is `SET' and the program accepts input of the format `SET data', the analysis tries to determine that `data' is a valid argument.
The arguments found are saved for later use.
Once every input pertaining to a permutation has been tested for arguments, the inputs are passed through a instrumented binary for the collection of execution trace.
The binary is instrumented with a Pin tool, developed by us, which logs the instructions executed as a function of time.
Once all permutations of a given size have been executed, the dynamic similarity test begins.
This test splits each execution trace into parts, each part being associated with an input string.
We refer to the execution trace of the input string `SET' as $\alpha(SET)$.
For all permutations of the same size with the same last input string, the $\alpha(last_cmd)$ are clustered.
Given a permutation $p$ of size $s$ let $\alpha(p_s)$ represent the execution trace of the command at index $s$ of the permutation (i.e. last command of the permutation).
It is natural to assume that given the correct (or expected) order the of input strings $1$ through $(i-1)$ of the permutation, the input $p_s$ experiences a different execution trace (or control flow path) compared to an incorrect permutation (assuming it is the same index in the permutation).
This information is used to determine, for every input string discovered, if there exist a permutations which affects its control flow path.
If such a permutation is discovered, it is added to the protocol state machine.

@@

\section{Evaluation and Results} \label{results}
\subsection{Finding Input, Input Size and Classifying Input}
\tool{} was tested on \numbinaries{} x86 binaries to measure its effectiveness from the DARPA Cyber Grand Challenge Example set \cite{darpacgc}.
The results are summarized in Figure \ref{fig:findinputresults}.
Analyzing the source code showed that out of the total \numbinaries{} binaries, 21 had predefined valid input strings.
\tool{} found input strings for 14 of the 21 binaries and categorize the input correctly for 100\% of the 14 binaries.
%TODO
Our method was able to determine the input length for XXX of the binaries.

\begin{figure}[h!]
\caption{The results of our input finding method. `N/A' signifies that the binary did not have expect any predefined input strings but on the contrary accepted any input.}
\centering
\includegraphics[width=0.5\textwidth]{findinputresults}
\label{fig:findinputresults}
\end{figure}
%To further show the advantage of this method, we built a smart fuzzer which uses our method to determine input and protocol for a given binary.
%We use the smart fuzzer and compare it to black box fuzzers symbolic execution engines with respect to code coverage over units of time.
%Using such a comparison allows the reader to determine which method is more efficient while also performing well.

@@

Additionally, our input finding method was tested on a binary used for technical interviews.
The binary was designed as a reverse engineering challenge for interviewers.
To pass the challenge, the interviewer has to crack the secret passcode to the program.
The blog post describes the manual process used to crack the password \cite{interviewbinary}.
The manual process is tedious and it includes methods such as disassembling the binary, stepping through the instructions using a debugger, and reverse engineering the logic of the program.
As described in the blog post, the cracking of the binary can take several hours and requires advanced skills and years of background knowledge.
\tool{} cracked the binary in a matter of minutes autonomously, proving itself as a useful tool not only finding valid input but also cracking passwords.

\subsection{Protocol State Machine Generation}
% TODO
The evaluation of the protocol state machine generator was also done using the DARPA Cyber Grand Challenge binaries.
However, out of the total of \numbinaries{} binaries, only [NUMBER] had an implicit protocol.
Our method was able to generate a correct protocol state machine for [NUMBER]\% of these binaries.

@@

\section{Future Work and Limitations} \label{futurework}
While the method presented in this paper has many advantages, it also has a few limitations.
For certain programs that allow any input, such as an echo server, our method is not any more efficient than a black box fuzzer.
Additionally, our method cannot handle cases in which input is composed of multiple inter-dependent fields, such as when the value of a field determines the length of another field and when binaries validate input fields out of order.
This is one reason why our method failed to detect input for some of the binaries tested.
Another reason of failure is binaries which accept non printable input (e.g. hex characters, binary, etc.).

In the future, we wish to improve our current method.
First, we would like to extend it to handle any type of input.
Furthermore, we would like to apply the input finding technique recursively.
It is easy to imagine programs for which the second input varies based on the value of the first input.
Also, as we've experienced during our evaluation, there exist programs which first request for a username, input which will not have any predefined valid values.
It would be helpful to determine such cases, provide some input and continue the execution of the program, applying the input finding technique once again.
However, there are challenges with this, such as determining when to not apply the finding input method because the program has reached an already observed and tested state.

With respect to the protocol state machine, currently our method uses the input finding technique to determine valid arguments.
However, the protocol state machine does not currently handle situations in which there are no predefined arguments (e.g. given input ``CREATE'' the arguments could be predefined such as ``dog'' or random such as ``x'').
Ultimately, we would like to generate some random arguments that remain constant for every input string in a given permutation.
Moreover, our current approach is not capable of identifying equivalent states. As seen in Figure \ref{fig:psm_example}, the different authentication paths (use of ``HELLO" vs. omission of ``HELLO") lead to different states when in reality the two states are identical.
Also, our method is based on the assumption that given a number of input strings, the majority of input string permutations are not valid and do not reflect the expected protocol.
There may be cases where this assumption does not hold.
In the future, we would like build onto our current approach for generating protocol state machines. 
First, we want to be able to determine if two states are identical for a more accurate protocol state machine.
Furthermore, we want to determine a better approach for identifying the valid sequence of input strings.
One option would be to monitor other aspects of the program, particularly the \textit{.data}, \textit{.bss}, and the heap during the executions.
Changes in these may reflect changes of the state of the program.

@@

We show that our proposed methods are efficient and successful by testing them on \numbinaries{} DARPA binaries from the Cyber Grand Challenge \cite{darpacgc}. 
Unlike previous works, our methods do not have any fundamental limitations (such as constraint solvers) and are robust enough to handle even binaries of other formats.
